A Large-Scale Study of the Use of Eval in R

### 2020-08-24 inst/analysis/corpus-analysis.Rmd
Regenerated data for V

```{r setup, include=FALSE}
for (x in c(
  "dplyr", 
  "DT",
  "fs",
  "fst",
  "ggplot2",
  "ggthemes",
  "lubridate",
  "readr",
  "purrr",
  "stringr",
  "tidyr"
  )) {
  suppressPackageStartupMessages(library(x, character.only=TRUE))
}

@@

```{r}
corpus <- as_tibble(read_fst(CORPUS_DETAILS_FILE))
```

## Corpus Overview

### Table

```{r summary}
overview_table(
  r("packages",             corpus),
  r("runnable code",        sum(corpus$runnable_code, na.rm=T)),
  r("examples code",        sum(corpus$runnable_code_examples, na.rm=T)),
  r("tests code",           sum(corpus$runnable_code_tests, na.rm=T)),
  r("vignettes code",       sum(corpus$runnable_code_vignettes , na.rm=T)),
  r("mean expr coverage",   percent(mean(corpus$coverage_expr))),
  r("median expr coverage", percent(median(corpus$coverage_expr))),
  r("with R code",          filter(corpus, package_r_code > 0)),
  r("R code",               sum(corpus$package_r_code, na.rm=T)),
  r("with native code",     filter(corpus, package_native_code > 0)),
  r("native code",          sum(corpus$package_native_code, na.rm=T)),
  r("functions",            sum(corpus$funs_public+corpus$funs_private, na.rm=T)),
  r("public functions",     sum(corpus$funs_public, na.rm=T)),
  r("private functions",    sum(corpus$funs_private, na.rm=T)),
  r("ranked in cranlogs",   filter(corpus, !is.na(rank))),
  r("mean revdes",          mean(corpus$revdeps)),
  r("median revdes",        median(corpus$revdeps)),
  r("runtime",              as.duration(sum(corpus$runnable_time))),
  r("mean runtime",         as.duration(mean(corpus$runnable_time))),
  r("median runtime",       as.duration(median(corpus$runnable_time))),
  r("function with eval",   sum(corpus$funs_with_eval, na.rm=T))
)
```

### CRAN

```{r}
overview_table(
  r("all cran", dim(available.packages())[1])
)
```

### Plots

@@

```{r, eval=FALSE}
parallel_log <- 
  read_csv(PARALLEL_LOG_FILE)

parallel_log %>%
  mutate(
    output=show_url(path(V1, "run.R.out"))
  ) %>%
  my_datatable(escape=FALSE)
```

>>>>

```{r setup, include=FALSE}
for (x in c(
  "dplyr", 
  "DT",
  "fs",
  "fst",
  "ggplot2",
  "ggthemes",
  "lubridate",
  "readr",
  "purrr",
  "stringr",
  "tidyr",
  "viridis"
  )) {
  suppressPackageStartupMessages(library(x, character.only=TRUE))
}

@@

```{r load corpus}
corpus <- 
  read_fst(CORPUS_DETAILS_FILE) %>%
  as_tibble() %>%
  mutate_at(
    vars(
      starts_with("runnable_code"),
      starts_with("runnable_files"),
      starts_with("package_"),
      starts_with("funs_"),
    ), 
    replace_na, 
    0
  )
```

```{r load kaggle}
kaggle_kernels <- as_tibble(read_fst(KAGGLE_KERNELS_FILE))
kaggle_log <- as_tibble(read_fst(KAGGLE_LOG_FILE)) %>%
  transmute(
    id=str_replace(Command, ".*/kaggle-run/[^/]+/([^/]+)/.*", "\\1"),
    exitval=Exitval,
    runtime=as.duration(JobRuntime)
  )

kaggle <- left_join(
  kaggle_kernels,
  kaggle_log,
  by="id"
)
```

```{r load runnable code metadata}
runnable_code_metadata <- read_fst(PACKAGE_RUNNABLE_CODE_FILE) %>%
  as_tibble() %>%
  mutate(package=basename(package)) %>%
  select(package, file=filename, code)

# TODO semi_join with traced-programs.fst
```

```{r load run metadata}
runs <- read_fst(PACKAGE_RUNS_FILE) %>%
  as_tibble() %>%
  mutate(
    package=basename(package),
    file=as.character(path("./", basename(dirname(file)), basename(file)))
  ) %>%
  semi_join(corpus, by="package")
```

## CRAN packages

```{r summary}
corpus_runtime  <- corpus$runnable_time / corpus$runnable_files
runnable_code   <- corpus$runnable_code
corpus_programs_files <- corpus$runnable_files

overview_table(
  r("packages",             corpus),
  r("in cranlogs",          filter(corpus, !is.na(rank))),
  r("package programs",     sum(corpus_programs_files)),

  r("runnable code",        sum(runnable_code)),
  r("runnable code mean",   mean(runnable_code)),
  r("runnable code median", median(runnable_code)),

  r("examples code",        sum(corpus$runnable_code_examples)),
  r("tests code",           sum(corpus$runnable_code_tests)),
  r("vignettes code",       sum(corpus$runnable_code_vignettes)),

  r("mean expr coverage",   percent(mean(corpus$coverage_expr))),
  r("median expr coverage", percent(median(corpus$coverage_expr))),

  r("with R code",          filter(corpus, package_r_code > 0)),
  r("with native code",     filter(corpus, package_native_code > 0)),

  r("code",                 sum(corpus$package_code)),
  r("R code",               sum(corpus$package_r_code)),
  r("native code",          sum(corpus$package_native_code)),

  r("functions",            sum(corpus$funs_public+corpus$funs_private)),
  r("function with eval",   sum(corpus$funs_with_eval)),
  r("public functions",     sum(corpus$funs_public)),
  r("private functions",    sum(corpus$funs_private)),

  r("mean revdes",          mean(corpus$revdeps)),
  r("median revdes",        median(corpus$revdeps)),

  r("runtime",              as.duration(sum(corpus$runnable_time))),
  r("mean runtime",         as.duration(mean(corpus_runtime))),
  r("median runtime",       as.duration(median(corpus_runtime)))
)
```

```{r generals}
overview_table(
  r("all cran", dim(available.packages())[1]),
  r("core packages", length(CORE_PACKAGES))
)
```

### Plot

@@

```{r}
kaggle_runnable <- filter(kaggle, runnable)
kaggle_finished <- filter(kaggle, exitval==0)
kaggle_code     <- kaggle_finished$code
kaggle_runtime  <- kaggle_finished$runtime

overview_table(
  r("kaggle",                 kaggle),
  r("runnable kaggle",        kaggle_runnable),
  r("finished kaggle",        kaggle_finished),
  r("finished kaggle code",   sum(kaggle_code)),
  r("finished kaggle mean",   mean(kaggle_code)),
  r("finished kaggle median", median(kaggle_code)),
  r("kaggle runtime",         as.duration(sum(kaggle_runtime))),
  r("kaggle mean",            as.duration(mean(kaggle_runtime))),
  r("kaggle median",          as.duration(median(kaggle_runtime)))
)
```

## Combined

```{r}
overview_table(
  r("all programs", nrow(kaggle_finished) + sum(corpus_programs_files)),
  r("all runnbale code", sum(kaggle_code) + sum(runnable_code))
)
```

```{r}
corpus_programs <-
  left_join(
    filter(runs, exitval==0) %>% select(package, file, time), 
    runnable_code_metadata %>% select(package, file, code),
    by=c("package", "file")
  ) %>%
  transmute(id=paste(package, ":", file), code, runtime=time, corpus="packages")

kaggle_programs <-
  kaggle %>%
  filter(exitval==0) %>%
  transmute(id, code, runtime=as.numeric(runtime, units="seconds"), corpus="kaggle")

programs <- bind_rows(corpus_programs, kaggle_programs)
```

```{r}
programs %>%
  ggplot(aes(y=code, x=corpus, fill=corpus)) + 
  geom_flat_violin(trim=FALSE, width=.5) +
  geom_dotplot(
    binaxis="y", 
    dotsize=0.1, 
    stackdir="down", 
    binwidth=0.01, 
    position=position_nudge(-0.025)) +
  scale_y_log10(labels = scales::comma) +
  scale_x_discrete(labels=c("kaggle"="Kaggle", "packages"="Packages")) +
  scale_fill_viridis(discrete = TRUE) +
  guides(fill = FALSE) +
  labs(y="Code size (log)") +
  theme(
    legend.box="horizontal",
    axis.title.x=element_blank()
  )

ggsave(path(PLOT_DIR, "programs-code.pdf"), width=7, height=5.5)
```

```{r}
programs %>%
  ggplot(aes(y=runtime, x=corpus, fill=corpus)) + 
  geom_flat_violin(trim=FALSE, width=.5) +
  geom_dotplot(
    binaxis="y", 
    dotsize=0.1, 
    stackdir="down", 
    binwidth=0.01, 
    position=position_nudge(-0.025)) +
  scale_y_log10(labels = function(x, ...) {
    min <- as.integer(x/60L)
    sec <- as.integer(x%%60L)
    sprintf("%.2d:%.2d", min, sec)
  }) +
  scale_x_discrete(labels=c("kaggle"="Kaggle", "packages"="Packages")) +
  scale_fill_viridis(discrete = TRUE) +
  guides(fill = FALSE) +
  labs(y="Running time [mm:ss] (log)") +
  theme(
    legend.box="horizontal",
    axis.title.x=element_blank()
  )
ggsave(path(PLOT_DIR, "programs-runtime.pdf"), width=3.5, height=5.0)
```

### 2020-08-24 inst/paper/main.tex
Add own, dependent and core eval calls from CRAN and Kaggle
Regenerated data for V

@@

@@

>>>>

We have assembled a corpus that consists of \CorpusPackages R packages and user programs.

@@

\begin{figure}[!tb]\centering\includegraphics[width=.9\linewidth]
  {programs-code.pdf}\caption{Corpus}\label{fig:corpus}
\end{figure}
\begin{figure}[!tb]\centering\includegraphics[width=.9\linewidth]
  {programs-runtime.pdf}\caption{Corpus}\label{fig:corpus}
\end{figure}

@@

\begin{table}[ht]
  \resizebox{\columnwidth}{!}{%
    \begin{tabular}{r|rrr|}
      \hline
      Source & Core & Dependent & Own \\
      \hline
      CRAN & \CranCoreTotalEvalCallsRnd (\CranCoreMinEvalCallsRnd/\CranCoreAvgEvalCallsRnd/\CranCoreMaxEvalCallsRnd) & \CranDependentTotalEvalCallsRnd (\CranDependentMinEvalCallsRnd/\CranDependentAvgEvalCallsRnd/\CranDependentMaxEvalCallsRnd) & \CranOwnTotalEvalCallsRnd (\CranOwnMinEvalCallsRnd/\CranOwnAvgEvalCallsRnd/\CranOwnMaxEvalCallsRnd) \\
      Kaggle & \KaggleCoreTotalEvalCallsRnd (\KaggleCoreMinEvalCallsRnd/\KaggleCoreAvgEvalCallsRnd/\KaggleCoreMaxEvalCallsRnd) & \KaggleDependentTotalEvalCallsRnd (\KaggleDependentMinEvalCallsRnd/\KaggleDependentAvgEvalCallsRnd/\KaggleDependentMaxEvalCallsRnd) & 0(0/0/0) \\
      \hline
    \end{tabular}
  }
  \label{table:source-eval-summary}
  \caption{Distribution of eval calls in CRAN and Kaggle}
\end{table}
